---
title: "Assemble Inputs"
author: "Lyndon Estes"
date: "January 20, 2016"
output: 
  html_document:
    highlight: tango
    theme: spacelab
    toc: yes 
    number_sections: yes
---

# Assemble Inputs

Bring all datasets in, create farmable area mask, mask them, and convert to data.tables for model input

1. For running model, non-grid: 
    + `currprod.rda` - For input, not data.table
    + `vprops.rda` - table of veg type area, rarity value, and prioritization value. 

2. For running model:
    + Prioritization (P) and impact (I)
        + `ZA-carbon.tif` - Carbon impacts, to data.table (P|I)
        + `bdprior.tif` - biodiversity prioritization grid (P)
        + `ZA-cost.tif` - cost distance
        + `maize|soybean_S|R|I_10km,tif` - variants thereof (P|I). Separate variants of these must be created prior to model runtime. A function will be creatd here to facilitate that.  
    + Impact only:
        + `ZA-vegremain.tif` - vegetation remaining grid, for assessing conversion impacts. 
    + Masking/allocation
        + `farmable-1k.tif` - slope-based farmable area mask
        + `water-1k.tif` - water-based farmable area mask
        + `ZA-remfrac.tif` - cropland and urban-based farmable area mask
3. For displaying model outputs: 
    + `cropland-1k.tif` - cropland mask

## Inputs
```{r, eval = FALSE}
library(agroEcoTradeoff)

p_proj <- "/Users/lestes/Dropbox/publications/zambiaToff/private/zambiaToff/"
p_devel <- full_path(p_proj, "external/input_devel/cropland")
p_basedat <- paste0("/Users/lestes/Dropbox/projects/agroEcoTradeoff/external/",
                    "base_data/")
p_crop <- full_path(p_proj, "external/input_devel/cropland")
p_bd <- full_path(p_proj, "external/input_devel/biodiversity")
p_cost <- full_path(p_proj, "external/input_devel/roads/cost")
p_yield <- full_path(p_proj, "external/input_devel/yield")
p_h20 <- full_path(p_proj, "external/input_devel/water")
p_terr <- full_path(p_proj, "external/input_devel/terrain")
p_carb <- full_path(p_proj, "external/input_devel/carbon")
p_production <- full_path(p_proj, "external/data/production")

# Load in data
mgrid <- raster(full_path(p_basedat, "ZA-grid.tif"))
cropland <- raster(full_path(p_crop, fname = "cropland-1k.tif"))
bd <- raster(full_path(p_bd, fname = "bdprior.tif"))
vegr <- brick(full_path(p_bd, fname = "ZA-vegremain.tif"))
load(full_path(p_bd, fname = "ZA-vegremain.rda"))
names(vegr) <- vegnames
carbon <- brick(full_path(p_carb, fname = "ZA-carbon.tif"))
names(carbon) <- c("veg", "soil")
cost <- raster(full_path(p_cost, fname = "ZA-cost.tif"))
ynms <- dir(p_yield, pattern = "tif")
ynms <- sapply(c("S", "R", "I"), function(x) ynms[grep(x, ynms)])
ynms <- ynms[c(1, 3, 5, 2, 4, 6)]
ylds <- lapply(ynms, function(x) {
  b <- brick(full_path(p_yield, x))
  ylnms <- gsub("\\.tif", "\\.rda", x)
  load(full_path(p_yield, ylnms))
  names(b) <- layernms
  b
})
names(ylds) <- gsub("_10K", "", ynms)

# mask materials
remfrac <- raster(full_path(p_bd, fname = "ZA-remfrac.tif"))
water <- raster(full_path(p_h20, fname = "water-1k.tif"))
farmable <- raster(full_path(p_terr, fname = "farmable-1k.tif"))

```

## Create farmable area

```{r, eval=FALSE}
# merge together the three farmland available masks
avail <- remfrac - water
avail[avail < 0] <- 0
avail <- avail - (1 - farmable)
avail[avail < 0] <- 0
avail <- avail * (bd < 0.962)  # remove PAs entirely (1 if nps, 0.962 if GMAs)
# plot(avail)
```

That creates the farmland availability mask that is needed.  

## Stack and mask all grids
```{r, eval = FALSE}

# first get rid of all NA area in all grids
# start with yield grids
yldmask <- calc(stack(ylds), sum)
yldmask2 <- disaggregate(yldmask, fact = 10)
yldmask3 <- resample(yldmask, yldmask2)
yldmask4 <- projectRaster(yldmask3, mgrid)
yldmask5 <- mask(yldmask4, mgrid)
plot(yldmask5)

# now with other
carbonf <- stack(lapply(1:nlayers(carbon), function(x) {
  r <- carbon[[x]]
  r[is.na(r)] <- 0
  r
})) 

mgridc <- crop(mgrid, avail)
inlist <- list("convertible" = avail, "vegremain" = vegr, "carbon" = carbonf, 
               "cost" = cost, "bd" = bd, "cropland" = cropland)
inlist <- lapply(inlist, function(x) crop(x, mgridc))
inlist2 <- lapply(inlist, function(x) calc(x, mean))
inlist3 <- calc(stack(inlist2), mean)

# filter NA values
nafilt <- calc(stack(!is.na(inlist3), !is.na(crop(yldmask5, mgridc))), sum)
# plot(nafilt)
# plot(!is.na(calc(vegr, mean)))
# plot(!is.na(crop(yldmask5, mgridc)), col = c("transparent", "red"), add = TRUE)
msk <- nafilt == maxValue(nafilt)
msk[msk == 0] <- NA

# updated mask
mgridf <- mask(mgridc, msk)
valinds <- which(!is.na(values(mgridf)))
outlist <- inlist[names(inlist) != "cropland"]

# write out new data.tables
cnames <- "ZA"  
onms <- names(inlist)
onms[onms == "bd"] <- "cons-priorities"
for(i in 1:length(outlist)) { # i <- 1
  fnm <- full_path(p_production, paste0(cnames, "-", onms[i], ".csv"))
  DT <- as.data.table(outlist[[i]], xy = FALSE)[valinds, ]
  write.table(DT, file = fnm, sep = ",", col.names = TRUE, row.names = FALSE)
}

# write out mask raster
base <- as.data.table(msk, xy = TRUE)
setnames(base, old = names(msk), new = "val")
base[, ind := 1:nrow(base)]
base <- base[!is.na(val)][, val := NULL]
fnm <- full_path(p_production, paste0(cnames, "-mask.csv"))
write.table(base, file = fnm, sep = ",", col.names = TRUE, row.names = FALSE)
# plot(dt_to_raster(base, mgridf@crs))

```

## Yield grids
```{r, eval = FALSE}
# Function to disaggregate and reproject yield rasters
yld_disagg <- function(x, msk, fact = 10) {
  r <- disaggregate(x, fact = 10)
  r2 <- resample(x, r)
  r3 <- projectRaster(r2, msk)
  r4 <- mask(r3, msk)
  r4
}

# Function to aggregate yields within a particular yield brick
# stat <- "50"; cults <- paste0("9900", c("08", "09", "10"))
yield_stat <- function(x, stat = "mu", mgmnt = "S", cults, fun = mean) {
  repstr <- paste0(paste0(stat, "_", mgmnt), cults, collapse = "|")
  cnms <- names(x)[grep(repstr, names(x))]
  print(cnms)
  calc(x[[cnms]], fun = fun)
}
# rm(stat, mgmnt, cults, fun, x, cnms, repstr)

# Let's make an initial yield combination for model development that is
# is composed of: 
# maize
# 1. Subsistence management, all four cultivars, median yields, 50%
# 2. Rainfed commerical, hybrid cultivars, median yields, 40%
# 3. Irrigated maize, hybrid cultivars, median yields, 10%

maize <- list("S" = yield_stat(x = ylds$maize_S.tif, "50", "S", 
                               paste0("000", 1:4)), 
              "R" = yield_stat(x = ylds$maize_R.tif, "50", "R", 
                               paste0("000", 3:4)), 
              "I" = yield_stat(x = ylds$maize_I.tif, "50", "I", 
                               paste0("000", 3:4)))
wgts <- c(0.5, 0.4, 0.1)
maizewgt <- lapply(1:length(wgts), function(x) maize[[x]] * wgts[x])
potmaize <- calc(stack(maizewgt), sum)  # potential maize yield

# Soybean
# 1. Subsistence management, all 3 cultivars, median yields, 20%
# 2. Rainfed commerical, all 3cultivars, median yields, 70%
# 3. Irrigated maize, all 3 cultivars, median yields, 10%
soy <- list("S" = yield_stat(x = ylds$soybean_S.tif, "50", "S", 
                             paste0("9900", c("08", "09", "10"))), 
            "R" = yield_stat(x = ylds$soybean_R.tif, "50", "R", 
                             paste0("9900", c("08", "09", "10"))), 
            "I" = yield_stat(x = ylds$soybean_I.tif, "50", "I", 
                             paste0("9900", c("08", "09", "10"))))
wgts <- c(0.2, 0.7, 0.1)
soywgt <- lapply(1:length(wgts), function(x) soy[[x]] * wgts[x])
potsoy <- calc(stack(soywgt), sum)  # potential maize yield
# plot(potmaize); cellStats(potmaize, mean)
# plot(potsoy); cellStats(potsoy, mean)
# plot(potmaize)
# plot(potsoy)

# Disaggregate potential yields, stack, and round
ylds_1k <- lapply(list(potmaize, potsoy), function(x) {
  yld_disagg(x, msk = mgridf)
})
ylds_s <- stack(ylds_1k) 
names(ylds_s) <- c("maize", "soy")
ylds_s <- round(ylds_s)  

# data.table them
cnames <- "ZA"  
fnm <- full_path(p_production, paste0(cnames, "-potential-yields.csv"))
DT <- as.data.table(ylds_s, xy = FALSE)[valinds, ]
write.table(DT, file = fnm, sep = ",", col.names = TRUE, row.names = FALSE)





```



