---
title: "Current Cropland, Production, and Yield"
author: "Lyndon Estes"
date: "January 14, 2016"
output: 
  html_document:
    highlight: tango
    theme: spacelab
    toc: yes 
    number_sections: yes
---

# New cropland map of Zambia

Cropland class from the SERVIR landcover map (2010)

## Inputs
```{r, eval = FALSE}
library(agroEcoTradeoff)
library(rworldxtra)
data("countriesHigh")

p_proj <- "/Users/lestes/Dropbox/publications/zambiaToff/private/zambiaToff/"
p_devel <- full_path(p_proj, "external/input_devel/cropland")
p_basedat <- paste0("/Users/lestes/Dropbox/projects/agroEcoTradeoff/external/",
                    "base_data/")
p_dat <- "~/Dropbox/data/agricultural/fao/countrystat/zambia/"
p_usda <- "~/Dropbox/data/agricultural/usda/fas/zambia/"
p_lc <- "~/Dropbox/data/landcover/Zambia_LandCover_2010_Scheme_II/"

# Load in data
mgrid <- raster(full_path(p_basedat, "ZA-grid.tif"))
lc <- raster(full_path(p_lc, fname = "Zambia_LandCover_2010_Scheme_II.tif"))

```

## Extract cropland class and resample
```{r, eval = FALSE}

# pullout cropland class 8
system.time(cropland <- rast_math(lc == 8, tmpdir = p_devel))

onm <- full_path(p_devel, "cropland-1k.tif")
ext <- bbox(mgrid)[1:4]  # extent   
gdalwarp(srcfile = cropland@file@name, t_srs = projection(mgrid), 
         dstfile = onm, r = "average", ot = "Float32", te = ext, 
         srcnodata = 255, dstnodata = 255, tr = c(1000, 1000), 
         of = "GTiff", multi = TRUE, verbose = FALSE, overwrite = TRUE)
cropland <- raster(onm)
par(mar = rep(0, 4)); plot(cropland)

```

## Cropland statistics

From FAO countrystat at this [link](http://www.countrystat.org/home.aspx?c=ZMB&tr=21).
```{r, eval = FALSE}
# select crop stats
fnms <- dir(p_dat, pattern = "10|16", full.names = TRUE)
vars <- c("prod", "pa")
fao <- rbindlist(lapply(1:length(fnms), function(x){
  DT <- fread(fnms[x], stringsAsFactors = FALSE)
  cbind(vars[x], DT)
})) 
fao[, Value := as.numeric(Value)]
setnames(fao, old = c("V1", "administrative level 1 (code)", 
                      "administrative level 1", "Crop (code)"), 
         new = c("var", "provcode", "prov", "cropcode"))
setkeyv(fao, c("var", "Year"))

# missing 2008 NW province maize and 2011 Southern Province, same for soybean
fao[Crop == "Maize" & var == "prod" & is.na(Value)]  
fao[Crop == "Soybeans" & var == "prod" & is.na(Value)]  

# corresponding FAOStat full country values are 1211566 and 3020380 for maize
# 56839 and 116539 for soybean
nwrepl <- fao[Crop == "Maize" & var == "prod" & Year == "2008" & 
                provcode != 3431, 1211566 - sum(Value)]
fao[Crop == "Maize" & var == "prod" & Year == "2008" & 
                provcode == 3431, Value := nwrepl]
srepl <- fao[Crop == "Maize" & var == "prod" & Year == "2011" & 
               provcode != 3433, 3020380 - sum(Value)]
fao[Crop == "Maize" & var == "prod" & Year == "2011" & 
                provcode == 3433, Value := srepl]
nwrepl <- fao[Crop == "Soybeans" & var == "prod" & Year == "2008" & 
                provcode != 3431, 56839 - sum(Value)]
fao[Crop == "Soybeans" & var == "prod" & Year == "2008" & 
                provcode == 3431, Value := nwrepl]
srepl <- fao[Crop == "Soybeans" & var == "prod" & Year == "2011" & 
               provcode != 3433, 116539 - sum(Value)]
fao[Crop == "Soybeans" & var == "prod" & Year == "2011" & 
                provcode == 3433, Value := srepl]
save(fao, file = full_path(p_devel, "fao-yields.rda"))

# Planted area and production 2010-2014 for Zambia
muprod <- lapply(c("Maize", "Soybeans"), function(x) {
  fao[Crop == x & var == "prod" & Year %in% c(2008:2015), sum(Value), 
      by = Year]
})
names(muprod) <- c("maize", "soy")
mupa <- lapply(c("Maize", "Soybeans"), function(x) {
  fao[Crop == x & var == "pa" & Year %in% c(2008:2015), sum(Value), 
      by = Year]
})
names(mupa) <- c("maize", "soy")
# mean(muprod$soy[, V1] / mupa$soy[, V1])
# mean(muprod$maize[, V1] / mupa$maize[, V1])

# Read in USDA harvested area and production to get difference in harvested area
# and planted area
fnms <- dir(p_usda, full.names = TRUE)
vars <- c("maize", "soy")
usda <- rbindlist(lapply(1:length(fnms), function(x){
  DT <- fread(fnms[x], stringsAsFactors = FALSE, header = TRUE)
  # cbind(vars[x], DT)
}))
cols <- names(usda)[4:9]
usda_ylds <- usda[var == "prod", cols, with = FALSE] / 
  usda[var == "harvarea", cols, with = FALSE]
save(usda, file = full_path(p_devel, "usda-yields.rda"))

# FAO references start year, it seems for maize, but soybean is correct year
yp <- unlist(usda[var == "prod" & crop == "maize", cols, with = FALSE] * 1000)
plot(yp, muprod[[1]]$V1[-c(1, 8)])  # but essentially same yields
yp <- unlist(usda[var == "prod" & crop == "soybean", cols, with =FALSE] * 1000)
plot(yp, muprod[[2]]$V1[-c(1:2)])  

# Production estimates ~2010/1-2015
currprod <- round(c("maize" = muprod$maize[Year %in% 2010:2014, mean(V1)], 
                    "soy" = muprod$soy[Year %in% 2011:2015, mean(V1)]))
save(currprod, file = full_path(p_devel, "currprod.rda"))

```

## Resulting model inputs

`currprod.rda`, `cropland-1k.tif`

Eventually might want to disaggregate yields onto cropland map, but will need provincial harvested area for that.  Model might not really need that, however. 

