---
title: "WB-IFAD presentation"
author: "Lyndon Estes"
date: "January 26, 2016"
output: 
  html_document:
    highlight: tango
    theme: spacelab
    toc: yes 
    number_sections: yes

---

# Initial slides for running model
```{r}
library(agroEcoTradeoff)
setwd("agroEcoTradeoff")
p_model <- set_base_path()
p_main <- "/Users/lestes/Dropbox/publications/zambiaToff/private/zambiaToff/"
p_figures <- full_path(p_main, "paper/figures/presentation/")
p_bd <- full_path(p_main, "external/input_devel/biodiversity")
library(rworldxtra)
data("countriesHigh")
# load other data here

```


## To make

### Meeting demand on existing land versus getting it all from new land

Existing yields are 2.3 t/ha for maize and 2 for soy.  Potential mean yields are 4.43 for maize and 3.4 for soybean, based on 2015-2050 projected changes from Bruinsma.  

Scenario: maize demand 2X, soy demand 5X

```{r, eval=FALSE}
setwd("agroEcoTradeoff")
path <- full_path(set_base_path(), "external/data/ZA")
code <- run_code("ZA")
il <- input_handler(path, ybetas = list(1, 1), ybeta_update = 0, code = code)

targ <- c("maize" = 3, "soy" = 6)
currprodmod <- c(1, 1)
target <- targets(prod_targ = targ, currprod = il$currprod, 
                  currprodmod = currprodmod)
mzgap <- 4.43 / 2.3
soygap <- 3.4 / 2
# 3224 / 1674; 2124 / 1257

# gap closing test
# y2 <- 7.5  # future yield
# y1 <- 5  # current yield
# ha <- 10  # hectares
# prod <- c(y2, y1) * ha  # future and current production
# (y1 + (2.5 * 0.25)) * ha  
# # future production with 25% of possible yield increase met
# (((prod[2] * y2 / y1) - prod[2]) * 0.25 + prod[2]) / prod[2] * prod[2]

# closing 75% of yield gap
cprodmod_h <- ((((il$currprod * c(mzgap, soygap)) - il$currprod) * 
                  1) + il$currprod) / il$currprod
# assume 75% gap for each
targ_high <- targets(targ, il$currprod, cprodmod_h)
# (1 + ((mzgap - 1) * 0.75) * mzgap) * il$currprod
# (il$currprod - 1 + ((mzgap - 1) * 0.75) * il$currprod)

# closing 25% of yield gap
cprodmod_l <- ((((il$currprod * c(mzgap, soygap)) - il$currprod) * 
                  0.25) + il$currprod) / il$currprod
# assume 75% gap for each
# targ_low <- targets(targ, il$currprod, cprodmod_l)
targ_low <- targets(targ, il$currprod, currprodmod = 1)

# Run model
cbetas <- c("Y" = 0.5, "C" = 0, "BD" = 0, "COST" = 0.5)
thigh <- tradeoff_mod(prod_targ = targ, cbetas = cbetas, 
                      ybetas = list(1, 1), exist_list = il, 
                      currprodmod = cprodmod_h, input_key = "ZA", 
                      silent = FALSE)
tlow <- tradeoff_mod(prod_targ = targ, cbetas = cbetas, 
                     ybetas = list(1, 1), exist_list = il, 
                     #currprodmod = cprodmod_l, input_key = "ZA", 
                     currprodmod = 1, input_key = "ZA", 
                     silent = FALSE)
thigh$impacts
tlow$impacts

CRSobj <- CRS(il$sp$crs)

# rasters
rhigh <- dt_to_raster(thigh$conv, CRSobj)
rhigh <- calc(rhigh * c(1, 2), sum)
rlow <- dt_to_raster(tlow$conv, CRSobj)
rlow <- calc(rlow * c(1, 2), sum)

zambia <- countriesHigh[countriesHigh@data$SOVEREIGNT == "Zambia", ]
zambia <- spTransform(zambia, CRSobj)
p_figs <- full_path(p_main, "paper/figures/presentation/")
# p_roads <- full_path(p_main, "external/input_devel/roads/tmdpfr_links.sqlite")
load(full_path(p_model, "external/data/ZA/parks_roads.rda"))
save(pas, roads, zambia, 
     file = full_path(p_model, "external/data/ZA/parks_roads.rda"))
# roads <- readOGR(p_roads, layer = "tmdpfr_links")
# pas <- raster(full_path(p_bd, "ZA-pas.tif"))
# pmask <- !is.na(pas)
# pmask[pmask == 0] <- NA

# impact plots
i1 <- tlow$impacts
imp_plot(dtr = rlow, impacts = i1, fnm = "low_closure.png", 
         ftitle = "Current Yield")
i2 <- thigh$impacts
imp_plot(dtr = rhigh, impacts = i2, fnm = "high_closure.png", 
         ftitle = "FAO-Projected Yield")
dr <- (rlow > 0) - (rhigh > 0)
plot(dr) #; freq(dr)
```

### Optimal for each parameter
```{r, eval=FALSE}
cbetas <- c("Y" = 1, "C" = 0, "BD" = 0, "COST" = 0)
Y <- tradeoff_mod(prod_targ = targ, cbetas = cbetas, ybetas = list(1, 1),
                  exist_list = il, currprodmod = 1, input_key = "ZA",
                  silent = FALSE)
cbetas <- c("Y" = 0, "C" = 1, "BD" = 0, "COST" = 0)
C <- tradeoff_mod(prod_targ = targ, cbetas = cbetas, ybetas = list(1, 1),
                  exist_list = il, currprodmod = 1, input_key = "ZA",
                  silent = FALSE)
cbetas <- c("Y" = 0, "C" = 0, "BD" = 1, "COST" = 0)
BD <- tradeoff_mod(prod_targ = targ, cbetas = cbetas, ybetas = list(1, 1),
                   exist_list = il, currprodmod = 1, input_key = "ZA",
                   silent = FALSE)
consr <- dt_list_to_raster(il$mask, list(il$cons), CRSobj)
cbetas <- c("Y" = 0, "C" = 0, "BD" = 0, "COST" = 1)
COST <- tradeoff_mod(prod_targ = targ, cbetas = cbetas, ybetas = list(1, 1),
                     exist_list = il, currprodmod = 1, 
                     input_key = "ZA", silent = FALSE)

Yr <- calc(dt_to_raster(Y$conv, CRSobj) * c(1, 2), sum)
Cr <- calc(dt_to_raster(C$conv, CRSobj) * c(1, 2), sum)
BDr <- calc(dt_to_raster(BD$conv, CRSobj) * c(1, 2), sum)
COSTr <- calc(dt_to_raster(COST$conv, CRSobj) * c(1, 2), sum)

cnms <- names(cbetas)
mnms <- paste(c("Yield", "Carbon Protection", "Biodiversity Protection", 
                "Cost Minimization"), "Prioritized")
for(i in 1:length(cnms)) { # i = 1
  print(cnms[i])
  i1 <- get(cnms[i])$impacts
  imp_plot(dtr = get(paste0(cnms[i], "r")), impacts = i1, 
           fnm = paste0(cnms[i], ".png"), ftitle = mnms[i])  
}

```

### Equal Weights
```{r, eval = FALSE}
cbetas <- c("Y" = 0.25, "C" = 0.25, "BD" = 0.25, "COST" = 0.25)
YBCD <- tradeoff_mod(prod_targ = targ, cbetas = cbetas, ybetas = list(1, 1),
                     exist_list = il, currprodmod = 1, input_key = "ZA",
                     silent = FALSE)
YBCDr <- calc(dt_to_raster(YBCD$conv, CRSobj) * c(1, 2), sum)
imp_plot(dtr = YBCDr, impacts = YBCD$impacts, 
         fnm = "YBCD.png", ftitle = "Ag 25%, C 25%, BIOD 25%, COST 25%")  

```



### Optimal expansion based on selected weights
```{r, eval = FALSE}
yblist <- list(yb1 <- c(1, 1))
step <- 0.1
# cnames <- c("Y", "C", "BD", "COST"); step = 0.05
cnames <- c("Y", "C", "BD", "COST"); step = 0.2
# cnames <- c("Y", "C", "BD"); step = 0.25#, "BD", "COST"); step = 0.2
pareto_steps(cnames, step)
Yv = "conv_area"; Cv = "tot_C"; BDv = "priority"; COSTv = "mu_cost"
Yst = sum; Cst = sum; BDst = mean; COSTst = mean
a <- Sys.time()
print(a)
pare <- pareto(cnames, step, targ, yblist = list(1, 1),
               input_key = "ZA", Yv, Cv, BDv, COSTv, Yst, Cst, BDst, COSTst,
               todisk = TRUE, silent = FALSE, ncl = 8, path = path)
Sys.time() - a  # 1.75 minutes
pare[, plot(Y, C)]

parms <- batch_params(list(c(1, 1)), list(targ), pareto_steps(cnames, step))

ott <- data.frame(pare)
ott$colors <- rainbow(length(ott$ind))
ott$ind <- seq(1, length(ott$ind))

odfs <- lapply(cnames, function(x) {
  if(x %in% colnames(ott)) {
    odf <- ott[order(ott[, x]), ]
  } else{ 
    odf <- NULL
  }
  odf
})

pvec <- cnames
tvec <- c("by Area", "by C", "by BIOD", "by Cost")
yvec <- c("Area km^2", "C Mt", "BIOD", "Time")
div <- c(1 / 100, 1 / 100000, 100, 1)
dfs <- which(sapply(odfs, is.data.frame))
png(full_path(p_figs, "pareto_weights.png"), height = 700, width = 1000)
if(length(dfs) == 2) {
  par(mfcol = c(2, 2), oma = c(2, 5, 1, 0), mar = c(1, 1, 1, 1) + 0.1, 
      xaxs = "i", mgp = c(1, 0.1, 0), tcl = -0.2)
} else if(length(dfs) == 3) {
  par(mfcol = c(3, 2), oma = c(2, 5, 1, 0), mar = c(1, 1, 1, 1) + 0.1, 
      xaxs = "i", mgp = c(1, 0.1, 0), tcl = -0.2)
} else if(length(dfs) == 4) {
  par(mfcol = c(4, 2), oma = c(2, 5, 1, 0), mar = c(1, 1, 1, 1) + 0.1, 
      xaxs = "i", mgp = c(1, 0.1, 0), tcl = -0.2)
}
for(i in 1:2) {  #i <- 1
  d <- odfs[[i]]
  if(!is.null(d)) {
    for(j in 1:length(pvec)) {  # j <- 1
      if(pvec[[j]] %in% colnames(d)) {
        barplot(round(d[, pvec[j]] * div[j]), names.arg = d$ind, col = d$colors,
                yaxt = "n", las = 2, cex.axis = 0.3)
        if(j == 1) mtext(tvec[i], side = 3)
        if(i == 1) mtext(yvec[j], side = 2, cex = 0.8, line = 3)
        if(i == 1) {
          axis(2, las = 2, cex = 0.6)
        } else {
          axis(2, labels = FALSE)
        }
      }
    } 
  }
}
dev.off()
quantile(0:100, probs = seq(0, 1, 0.1))

# Figure out which meets multiple objective simultaneously
pdeciles <- pare[, lapply(.SD, function(x) {
  quantile(x, probs = seq(0, 1, 0.1))[-1]
  }), .SD = 1:4]
# pdeciles <- cbind("deciles" = paste0(seq(10, 100, 10), "th"), pdeciles)

prange <- pare[, lapply(.SD, range), .SD = 1:4]
pdiff <- prange[, lapply(.SD, function(x) diff(x) * 0.4 + min(x))]
swvec <- as.matrix(pdiff)[1, ]
tmat <- sweep(as.matrix(pare)[, 1:4], MARGIN = 2, swvec, `<=`)
tmat[tmat == "TRUE"] <- 1

best_ind <- which(rowSums(tmat[, 1:4]) == 4)
parms[best_ind, 1:4]
pare[best_ind, ]

# alternative scenario
pdiff2 <- prange[, lapply(.SD, function(x) diff(x) * 0.2 + min(x))]
pdiff3 <- prange[, lapply(.SD, function(x) diff(x) * 0.6 + min(x))]
pdiff4 <- cbind(pdiff2[, 1:2, with = FALSE], pdiff3[, 3:4, with = FALSE])

swvec <- as.matrix(pdiff4)[1, ]
tmat <- sweep(as.matrix(pare)[, 1:4], MARGIN = 2, swvec, `<=`)
tmat[tmat == "TRUE"] <- 1
ind1 <- which(rowSums(tmat[, 1:4]) == 4)  # no solution

swvec <- as.matrix(pdiff2)[1, 1:2]
tmat <- sweep(as.matrix(pare)[, 1:2], MARGIN = 2, swvec, `<=`)
tmat[tmat == "TRUE"] <- 1
second_ind <- which(rowSums(tmat[, 1:2]) == 2)

parms[second_ind, 1:4]
pare[best_ind, ]


# sapply(1:nrow(pdeciles) {
#   #swvec <- as.matrix(pdeciles[deciles == "20th", cnames, with = FALSE])[1, ]
#   swvec <- as.matrix(pdeciles[deciles == "20th", cnames, with = FALSE])[1, ]
#   tmat <- sweep(as.matrix(pare)[, 1:4], MARGIN = 2, swvec, `<=`)
#   parms
# })
# pare[, lapply(.SD, function(x) )]

```

### Create map of those weight combinations
```{r, eval = FALSE}
cbetas <- parms[best_ind, 1:4]
tobest <- tradeoff_mod(prod_targ = targ, cbetas = cbetas, 
                       ybetas = list(1, 1), exist_list = il, 
                       currprodmod = 1, input_key = "ZA", silent = FALSE)
rbest <- calc(dt_to_raster(tobest$conv, CRSobj) * c(1, 2), sum)
ib <- tobest$impacts
imp_plot(dtr = rbest, impacts = ib, fnm = "preferred_scenario.png", 
         ftitle = "Ag 40%, C 0%, BIOD 40%, COST 20%")

cbetas <- parms[second_ind, 1:4]
tosecond <- tradeoff_mod(prod_targ = targ, cbetas = cbetas, 
                         ybetas = list(1, 1), exist_list = il, 
                         currprodmod = 1, input_key = "ZA", 
                         silent = FALSE)
rsecond <- calc(dt_to_raster(tosecond$conv, CRSobj) * c(1, 2), sum)
i2b <- tosecond$impacts
imp_plot(dtr = rsecond, impacts = i2b, fnm = "second_scenario.png", 
         ftitle = "Ag 60%, C 40%, BIOD 0%, COST 0%")

```

### Heat map from all combinations
```{r, eval = FALSE}
bpath <- full_path(p_model, "external/output/batch/dt/")
load(full_path(bpath, "bcode.rda"))
fnames <- dir(full_path(bpath, bcode), pattern = "csv", full.names = TRUE)
DT <- lapply(fnames, function(x) fread(x)[, rowSums(.SD)])
DTbind <- data.table(do.call(cbind, DT))
DTsum <- data.table(DTbind[, rowSums(.SD)])
DTspat <- cbind(il$mask[, .(x, y)], DTsum)
DTb <- dt_to_raster(DTspat, CRSobj)
DTbpct <- DTb / maxValue(DTb) * 100

DTbpct[DTbpct == 0] <- NA

png(full_path(p_figs, "heat.png"), height = 700, width = 700)
ccols <- c("grey90", "orange", "red")
pacols <- c("transparent", rgb(1, 0, 0, alpha = 0.1),
            rgb(0, 0, 1, alpha = 0.1), "transparent")
par(mar = c(1, 1, 1, 3))
plot(zambia, col = "grey90", border = "transparent")
plot(DTbpct, col = rev(heat.colors(20)), add = TRUE)
# plot(dtr, legend = FALSE, col = ccols, add = TRUE)
plot(pas[pas@data$type == "gma", ], add = TRUE, col = pacols[2],
     border = FALSE)
plot(pas[pas@data$type == "npark", ], add = TRUE, col = pacols[3],
     border = FALSE)
plot(roads, add = TRUE, lwd = 0.1, col = "grey")
dev.off()

```

### Slide showing an area equal to 200% of need
```{r, eval = FALSE}
cbetas <- parms[best_ind, 1:4]
targ200 <- c("maize" = 6, "soy" = 12)
to200 <- tradeoff_mod(prod_targ = targ200, cbetas = cbetas, 
                      ybetas = list(1, 1), exist_list = il, 
                      currprodmod = 1, input_key = "ZA", silent = FALSE)
r200 <- calc(dt_to_raster(to200$conv, CRSobj) * c(1, 2), sum)
ib <- to200$impacts
imp_plot(dtr = r200, impacts = ib, fnm = "demand2X.png", 
         ftitle = "Maize 6X, Soy 12X")
```

### Our areas versus proposed farm blocks
```{r, eval = FALSE}
# fb <- "~/Dropbox/projects/agroEcoTradeoff/external/base_data/farm-blocks.tif"
# fblocks <- raster(fb)
fbp <- "/Users/lestes/Dropbox/data/zari/misc/NEW4GEO/Fblocks.sqlite"
fb <- readOGR(fbp, layer = "fblocks")
fbalb <- spTransform(fb, CRSobj)
fbalb$ID <- 1:nrow(fbalb)
fbalb@data <- fbalb@data[, "ID", drop = FALSE]
fnm <- full_path(p_main, "external/input_devel/other/fblocks_alb.sqlite")
writeOGR(fbalb, dsn = fnm, layer = "fblocks_alb", driver = "SQLite")
# First rasterize farm blocks
# require(gdalUtils)
# onm <- full_path(p_main, "external/input_devel/other/fblocks_alb.tif")
# gdal_rasterize(src_datasource = fnm, l = "fblocks_alb", ot = "Byte", 
#                a = "ID", tr = res(r200), at = TRUE, co = "COMPRESS=LZW",
#                dst_filename = fnm, te = bbox(r200)[1:4], of = "GTiff", 
#                verbose = TRUE)
fbr <- rasterize(fbalb, r200)
fbr[is.na(fbr)] <- 0
fbr[fbr > 0] <- 1
fbr <- mask(fbr, r200)

r200sum <- r200 > 0
fbint <- (r200sum == 1 & fbr == 1) * 3 + (r200sum == 1 & fbr == 0) * 2 +
  (r200sum == 0 & fbr == 1) * 1
png(full_path(p_figs, "fblock-overlap.png"), height = 700, width = 700)
ccols <- c("grey90", "tan", "green3", "red")
pacols <- c("transparent", rgb(1, 0, 0, alpha = 0.1),
            rgb(0, 0, 1, alpha = 0.1), "transparent")
par(mar = c(1, 1, 1, 3))
plot(zambia, col = "grey90", border = "transparent")
plot(fbint, col = ccols, add = TRUE, legend = FALSE)
# plot(dtr, legend = FALSE, col = ccols, add = TRUE)
plot(pas[pas@data$type == "gma", ], add = TRUE, col = pacols[2],
     border = FALSE)
plot(pas[pas@data$type == "npark", ], add = TRUE, col = pacols[3],
     border = FALSE)
plot(roads, add = TRUE, lwd = 0.1, col = "grey")
legend("bottomright", 
       legend = c("Farm blocks", "Converted", "Converted in farm blocks", 
                  "NatParks", "GMAs"),
       pch = 15, col = c(ccols[2:4], pacols[2:3]), bty = "n", pt.cex = 3,
       cex = 1.5)
dev.off()


```

### Slide showing clustering

### Slide showing clustering along proposed road improvement areas

### Remaining Natural Vegetation Types

```{r, eval = FALSE}
p_bd <- full_path(p_main, "external/input_devel/biodiversity")
remfrac <- raster(full_path(p_bd, "ZA-remfrac.tif"))
png(full_path(p_figs, "remveg.png"), height = 700, width = 700)
par(mar = c(1, 1, 1, 1))
plot(remfrac, axes = FALSE, box = FALSE)
dev.off()

```

### Background Slides
```{r, eval = FALSE}
brasts <- dt_list_to_raster(il$mask, 
                            il[c("p_yield", "carbon", "cons", "cost")], 
                            CRSobj)
brasts <- lapply(brasts, function(x) x[[-c(1:2)]])

# Yields
png(full_path(p_figs, "yields.png"), height = 350, width = 700)
par(mfrow = c(1, 2), mar = c(0, 0, 0, 3), oma = c(0, 0, 0, 2))
for(i in 1:2) {
  plot(zambia, col = "grey90", border = "transparent")
  plot(brasts$p_yield[[i]], add = TRUE)
  plot(pas[pas@data$type == "gma", ], add = TRUE, col = pacols[2], border = F)
  plot(pas[pas@data$type == "npark", ], add = TRUE, col = pacols[3], border = F)
  plot(roads, add = TRUE, lwd = 0.1, col = "grey")
  if(i == 2) {
    legend("bottomright", legend = c("NatParks", "GMAs"),
           pch = 15, col = pacols[2:3], bty = "n", pt.cex = 3,
           cex = 1.5)
  }
}
dev.off()

# Carbon loss
closs <- brasts$carbon[[1]] + brasts$carbon[[1]] * 0.25
png(full_path(p_figs, "pot-C-loss.png"), height = 700, width = 700)
par(mar = c(1, 1, 1, 3))
plot(closs, axes = FALSE, box = FALSE)
plot(pas[pas@data$type == "gma", ], add = TRUE, col = pacols[2], border = F)
plot(pas[pas@data$type == "npark", ], add = TRUE, col = pacols[3], border = F)
plot(roads, add = TRUE, lwd = 0.1, col = "grey")
legend("bottomright", legend = c("NatParks", "GMAs"),
       pch = 15, col = pacols[2:3], bty = "n", pt.cex = 3,
       cex = 1.5)
dev.off()

# Cost
png(full_path(p_figs, "cost.png"), height = 700, width = 700)
par(mar = c(1, 1, 1, 3))
plot(brasts$cost, axes = FALSE, box = FALSE)
plot(pas[pas@data$type == "gma", ], add = TRUE, col = pacols[2], border = F)
plot(pas[pas@data$type == "npark", ], add = TRUE, col = pacols[3], border = F)
plot(roads, add = TRUE, lwd = 0.1, col = "grey")
legend("bottomright", legend = c("NatParks", "GMAs"),
       pch = 15, col = pacols[2:3], bty = "n", pt.cex = 3,
       cex = 1.5)
dev.off()

# Biodiversity
png(full_path(p_figs, "biodiversity.png"), height = 700, width = 700)
par(mar = c(1, 1, 1, 3))
plot(brasts$cons, axes = FALSE, box = FALSE)
plot(pas[pas@data$type == "gma", ], add = TRUE, col = pacols[2], border = F)
plot(pas[pas@data$type == "npark", ], add = TRUE, col = pacols[3], border = F)
plot(roads, add = TRUE, lwd = 0.1, col = "grey")
legend("bottomright", legend = c("NatParks", "GMAs"),
       pch = 15, col = pacols[2:3], bty = "n", pt.cex = 3,
       cex = 1.5)
dev.off()

```



